---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
options(width = 80)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "Readme_files/"
)

## Helper to determine OPAL version:
v_major  = seq(4L, 10L)
v_minor1 = seq_len(10L) - 1
v_minor2 = seq_len(10L) - 1

versions = expand.grid(v_major, v_minor1, v_minor2)
od = order(versions[, 1], versions[, 2], versions[, 3])
versions = versions[od, ]
vstrings = paste(versions[, 1], versions[, 2], versions[, 3], sep = ".")

getOPALVersion = function(opal, versions) {
  k = 1
  ov = opalr::opal.version_compare(opal, versions[k])
  while (ov != 0) {
    if (ov > 0) k = k + 1
    if (ov < 0) stop("Version is smaller than the smallest one from vector.")
    ov = opalr::opal.version_compare(opal, versions[k])
  }
  return(versions[k])
}

pkgs = c("here", "opalr", "DSI", "DSOpal", "dsBaseClient")
for (pkg in pkgs) {
  if (! requireNamespace(pkg, quietly = TRUE))
    install.packages(pkg, repos = c(getOption("repos"), "https://cran.obiba.org"))
}
devtools::install(quiet = TRUE, upgrade = "always")

## Install packages on the DataSHIELD test machine:
surl     = "https://opal-demo.obiba.org/"
username = "administrator"
password = "password"

opal = opalr::opal.login(username = username, password = password, url = surl)
opal_version = getOPALVersion(opal, vstrings)

check1 = opalr::dsadmin.install_github_package(opal = opal, pkg = "dsBinVal", username = "difuture-lmu", ref = "main")
if (! check1)
  stop("[", Sys.time(), "] Was not able to install dsBinVal!")

check2 = opalr::dsadmin.publish_package(opal = opal, pkg = "dsBinVal")
if (! check2)
  stop("[", Sys.time(), "] Was not able to publish methods of dsBinVal!")

opalr::opal.logout(opal)
```
[![R-CMD-check](https://github.com/difuture-lmu/dsBinVal/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/difuture-lmu/dsBinVal/actions/workflows/R-CMD-check.yaml) [![License: LGPL v3](https://img.shields.io/badge/License-LGPL%20v3-blue.svg)](https://www.gnu.org/licenses/lgpl-3.0) [![codecov](https://codecov.io/gh/difuture-lmu/dsBinVal/branch/main/graph/badge.svg?token=E8AZRM6XJX)](https://codecov.io/gh/difuture-lmu/dsBinVal)

# ROC-GLM and Calibration for DataSHIELD

The package provides functionality to conduct and visualize ROC analysis and calibration on decentralized data. The basis is the [DataSHIELD](https://www.datashield.org/) infrastructure for distributed computing. This package provides the calculation of the [__ROC-GLM__](https://www.jstor.org/stable/2676973?seq=1) with [__AUC confidence intervals__](https://www.jstor.org/stable/2531595?seq=1) as well as calibration curves and the brier score. In order to calculate the ROC-GLM or assess calibration it is necessary to push models and predict them at the servers which is also provided by this package. Note that DataSHIELD uses an option `datashield.privacyLevel` to indicate the minimal amount of numbers required to be allowed to share an aggregated value of these numbers. Instead of setting the option, we directly retrieve the privacy level from the [`DESCRIPTION`](https://github.com/difuture-lmu/dsBinVal/blob/master/DESCRIPTION) file each time a function calls for it. This options is set to 5 by default. The methodological base of th epackage is explained in detail [here](https://arxiv.org/abs/2203.10828).

## Installation

At the moment, there is no CRAN version available. Install the development version from GitHub:

```{r,eval=FALSE}
remotes::install_github("difuture-lmu/dsBinVal")
```

#### Register methods

It is necessary to register the assign and aggregate methods in the OPAL administration. These methods are registered automatically when publishing the package on OPAL (see [`DESCRIPTION`](https://github.com/difuture-lmu/dsBinVal/blob/main/DESCRIPTION)).

Note that the package needs to be installed at both locations, the server and the analysts machine.

## Installation on DataSHIELD

The two options are to use the Opal API:

- Log into Opal ans switch to the `Administration/DataSHIELD/` tab
- Click the `Add DataSHIELD package` button
- Select `GitHub` as source, and use `difuture-lmu` as user, `dsBinVal` as name, and `main` as Git reference.

The second option is to use the `opalr` package to install `dsBinVal` directly from `R`:
```{r, eval=FALSE}
### User credentials (here from the opal test server):
surl     = "https://opal-demo.obiba.org/"
username = "administrator"
password = "password"

### Install package and publish methods:
opal = opalr::opal.login(username = username, password = password, url = surl)

opalr::dsadmin.install_github_package(opal = opal, pkg = "dsBinVal", username = "difuture-lmu", ref = "main")
opalr::dsadmin.publish_package(opal = opal, pkg = "dsBinVal")

opalr::opal.logout(opal)
```

## Usage

A more sophisticated example is available [here](github.com/difuture-lmu/datashield-roc-glm-demo).

```{r}
library(DSI)
library(DSOpal)
library(dsBaseClient)

library(dsBinVal)
```

#### Log into DataSHIELD server

```{r}
builder = newDSLoginBuilder()

surl     = "https://opal-demo.obiba.org/"
username = "administrator"
password = "password"

builder$append(
  server   = "ds1",
  url      = surl,
  user     = username,
  password = password
)
builder$append(
  server   = "ds2",
  url      = surl,
  user     = username,
  password = password
)

connections = datashield.login(logins = builder$build(), assign = TRUE)
```

#### Assign iris and validation vector at DataSHIELD (just for testing)

```{r}
datashield.assign(connections, "iris", quote(iris))
vcall = paste0("quote(c(", paste(rep(c(1, 0), times = c(50, 100)), collapse = ", "), "))")
datashield.assign(connections, "y", eval(parse(text = vcall)))
```

#### Load test model, push to DataSHIELD, and calculate predictions

```{r}
# Model predicts if species of iris is setosa or not.
iris$y = ifelse(iris$Species == "setosa", 1, 0)
mod = glm(y ~ Sepal.Length, data = iris, family = binomial())

# Push the model to the DataSHIELD servers using `dsPredictBase`:
pushObject(connections, mod)

# Calculate scores and save at the servers using `dsPredictBase`:
pfun =  "predict(mod, newdata = D, type = 'response')"
predictModel(connections, mod, "pred", "iris", predict_fun = pfun)

datashield.symbols(connections)
```

#### Calculate l2-sensitivity

```{r}
# In order to securely calculate the ROC-GLM, we have to assess the
# l2-sensitivity to set the privacy parameters of differential
# privacy adequately:
l2s = dsL2Sens(connections, "iris", "pred")
l2s

# Due to the results presented in https://arxiv.org/abs/2203.10828, we set the privacy parameters to
# - epsilon = 0.2, delta = 0.1 if        l2s <= 0.01
# - epsilon = 0.3, delta = 0.4 if 0.01 < l2s <= 0.03
# - epsilon = 0.5, delta = 0.3 if 0.03 < l2s <= 0.05
# - epsilon = 0.5, delta = 0.5 if 0.05 < l2s <= 0.07
# - epsilon = 0.5, delta = 0.5 if 0.07 < l2s BUT results may be not good!
```

#### Calculate ROC-GLM

```{r}
roc_glm = dsROCGLM(connections, truth_name = "y", pred_name = "pred",
  dat_name = "iris", seed_object = "y")
roc_glm

plot(roc_glm)
```

#### Assess calibration

```{r}
dsBrierScore(connections, "y", "pred")

### Calculate and plot calibration curve:
cc = dsCalibrationCurve(connections, "y", "pred", 10, 3)
cc

plot(cc)
```

#### Further performance metrics

```{r}
dsConfusion(connections, "y", "pred")
```

## Deploy information:

__Build by `r Sys.info()[["login"]]` (`r Sys.info()[["sysname"]]`) on `r as.character(Sys.time())`.__

This readme is built automatically after each push to the repository. Hence, it also is a test if the functionality of the package works also on the DataSHIELD servers. We also test these functionality in `tests/testthat/test_on_active_server.R`. The system information of the local and remote servers are as followed:


```{r, include=FALSE}
ri_l  = sessionInfo()
ri_ds = datashield.aggregate(connections, quote(getDataSHIELDInfo()))
client_pkgs = c("DSI", "DSOpal", "dsBaseClient", "dsBinVal")
remote_pkgs = c("dsBase", "resourcer", "dsBinVal")
```

- Local machine:
    - `R` version: `r ri_l$R.version$version.string`
    - Version of DataSHELD client packages:


```{r, echo=FALSE}
dfv = installed.packages()[client_pkgs, ]
dfv = data.frame(Package = rownames(dfv), Version = unname(dfv[, "Version"]))
knitr::kable(dfv)
```

- Remote DataSHIELD machines:
    - `R` version of `r names(ri_ds)[1]`: `r ri_ds[[1]]$session$R.version$version.string`
    - `R` version of `r names(ri_ds)[2]`: `r ri_ds[[2]]$session$R.version$version.string`
    - OPAL version of the test instance: `r opal_version`
    - Version of server packages:


```{r, echo=FALSE}
dfv = do.call(cbind, lapply(names(ri_ds), function(nm) {
  out = ri_ds[[nm]]$pcks[remote_pkgs, "Version", drop = FALSE]
  colnames(out) = paste0(nm, ": ", colnames(out))
  as.data.frame(out)
}))
dfv = cbind(Package = rownames(dfv), dfv)
rownames(dfv) = NULL
knitr::kable(dfv)
```

```{r, include=FALSE}
datashield.logout(connections)
```

